---
title: "Are Public Libraries Become Less Popular? An Analysis of the Visits Number of Toronto Libraries over the Past Decade"
author: "Xu Qi"
thanks: "Code and data from this analysis are available at: https://github.com/xuqi2002/toronto_library_visit" 
date: "today"
date-format: "long" 
abstract: "Public libraries have  been vital community centers, offering a wealth of knowledge and resources. However, the advent of digital technology and the COVID-19 pandemic, which encouraged home-based activities, have impacted their popularity. An analysis of the Toronto Public Library's visitation data from 2012 to 2022 reveals a general decline in visits, with a significant drop during the pandemic. Despite some recovery post-pandemic, the decreased overall library visits suggest a reduced popularity, though this does not negate the ongoing importance of public libraries in society."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up ####

## Load packages
library(tidyverse)
library(knitr)
library(ggplot2)
library(dplyr)
library(opendatatoronto)

# Read in cleaned data #
cleaned_toronto_libraries <- read.csv(here::here("inputs/data/cleaned_toronto_libraries.csv"))
show_col_types = FALSE
```

## Introduction

Over the past decade, public libraries have been greatly affected by a number of factors. This is particularly evident in big cities like Toronto. This paper aims to analyze the trends in visit numbers to Toronto's public libraries over the past decade, providing insights into the popularity of these institutions.

In the digital age, where information can be easily accessed online. The popularity of digital resources such as e-books and online databases has dramatically changed the public's information access behavior [@rlu2023people]. In addition, the COVID-19 outbreak has caused people to spend more time at home [@rmcmenemy2023impact]. This study aims to understand whether these external factors affect the visit number in Toronto Public Library.

To examine the trend of visit number of Toronto Public Library, this paper is organized into the following sections: Data and Discussion. In the Data section, I introduced the nature of the data set obtained through the City of Toronto's OpenDataToronto Library [@rOpenDataToronto] and the procedures I took to clean and analyze the data. Trends identified during the analysis were also highlighted. In the discussion section, in addition to summarizing the findings during the data analysis process, the shortcomings of this study and the many aspects that can be improved to make this analysis more reliable are discussed.

## Data

The data set used in this paper was obtained through the City of Toronto's OpenDataToronto Library [@rOpenDataToronto]. It is entitled 'Library Visits' [@LibraryStatistics]. Data was collected and analyzed using the statistical programming software R [@r], with additional support packages including `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rcitekableExtra] and `knitr` [@rknitr]. Detailed data collection, cleaning and analysis process are below.

### Data collection

This dataset was published by the Toronto Public Library [@LibraryStatistics], provides an overview of number of annual visits to all branches of the Toronto Public Library from 2012 to 2022. The format of this dataset is table. Each row of the table has the year, and the corresponding number of visits, @tbl-cleaned_toronto_libraries is an example of first 5 rows of the cleaned data set. This dataset was last refreshed on June 30, 2023 and captured for this paper on January 21, 2024.

```{r}
#| message: false
#| echo: false
#| label: tbl-cleaned_toronto_libraries
#| tbl-cap: Sample of Cleaned Toronto Libraries Data

# Showing the first 5 rows of cleaned data
head (cleaned_toronto_libraries, 5) |>
  kable(
    col.names = c("Year", "Visits"),
  booktabs = TRUE
  )
```

### Data analysis

Based on the cleaned data, a line graph could show the trend of visit over years.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total Annual Visits at All Toronto Public Library Branches (2012-2022)", fig.width=7, fig.height=3.2}
#| label: fig-1
# Plotting the graph using ggplot2
ggplot(cleaned_toronto_libraries, aes(x = year, y = visits)) +
  geom_line(color = "blue", linewidth = 1) +  
  geom_point(color = "red", size = 2) +
  scale_x_continuous(breaks = seq(min(cleaned_toronto_libraries$year), max(cleaned_toronto_libraries$year), by = 1)) +
  labs(x = "Year",
       y = "Total Visits") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 5))
```

@fig-1 shows a gradual downward trend in total visits to the Toronto Public Library from 2012 to 2019. It declines even more sharply thereafter in 2020 and 2021. While it recovers in 2022, it is still only half of where it was 10 years ago. This trend is explained in the discussion part below with more details.

## Discussion

Analysis of the dataset shows a gradual decline in visits to the Toronto Public Library from 2012 to 2019. In 2020, the worst year of COVID-19, due to lockdown measures as well as health reasons, there is a very significant drop in the number of visits and the trend continues to 2021. By 2022, due to the better situation of COVID-19 and the gradual lifting of various measures, the number of visits has rebounded significantly, but it is still much lower than the year before 2020.

However, this analysis is not sufficient to show a very significant downward trend in the popularity of public libraries in Toronto. The first reason is that this analysis only considers data related to the number of visits, and therefore no sufficient evidence for specific reasons behind the decline I explained above. And the other reason is that data for 2023 is not included, due to the fact that the Toronto Public Library does not publish whole year data, yet. 2023 is the first year of the full deregulation of COVID-19, so it is not necessary for its data to be similar to years from 2020 to 2022, making it impossible to conclude the most recent trend.

In conclusion, this paper reveals the fact that the number of visits to the Toronto Public Library has been gradually decreasing over the past decade, but the reasons behind it and the trend after COVID-19 still need to be confirmed by analyzing more types and amounts of data.

## References
